---
title: "gun control chi-squared"
author: "Kirby Bartlett"
date: "November 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This: Kirby is updating markdown file here.
```{r, include=FALSE}
#load relevant libraries and data 
library(tidyverse)
gc_df <- read_csv("gc_df.csv")

```

Kirby: Is there a significant difference in reccommended gun control action for male and female respondents (should guns be easier, harder, or the same difficulty to purchase)?

Michael: test change 1

Kirby: data wrangling and counts 

```{r}
#counting how many males and females said about the three gun control options 

get_guns <- gc_df %>% 
  filter(gender != "NA", make_get_guns != "NA") %>% 
  count(gender, make_get_guns) %>%
  spread(make_get_guns, n) %>% # spread the data out based on the value of make_get_guns
  select(-gender)

rownames(get_guns) <- c("Female", "Male")

get_guns
```


My name is Michael the collaborator and I am going to be doing a chi-square test for independence.

Is there a significant association between sex and opinion on whether guns should be more difficult, the same, or easier to get?

```{r}

# lets look at the actual proportions 

guns_prop <- prop.table(as.matrix(get_guns), 1) # proportion calculated by row
guns_prop

# run a chi-square test for independence
guns_x2 <- chisq.test(get_guns)
guns_x2
```
Recommended government action on gun access differed significantly between male (n = ) and female (n = ) respondents ($\chi^2$(`r guns_x2$parameter`) = ..., p < 0.001). Notably, a much higher propotion of female respondents (57.4%) recommend making it more difficult to get guns, compared to 40.4% of males. 


